---
title: "Inferential Statistics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction to Inferential Statistics

> infer | ɪnˈfəː |
verb (infers, inferring, inferred) [with object]
deduce or conclude (something) from evidence and reasoning rather than from explicit statements: [with clause] : from these facts we can infer that crime has been increasing.

## Gaussian Distribution
The probability density function of the Gaussian distribution with mean $\mu$ and standard deviation $\sigma$ is given by
$$
P(x) = \frac{1}{\sqrt{2 π \sigma^2}} e^{-\frac{1}{2} (\frac{x - \mu}{\sigma})^{2}}
$$
```{r}
x <- seq(-10, 10, length = 1000)
y <- dnorm(x)
plot(x, y, type = "l")
```

Cumulative density function;
$$
\phi(x) = \int_{-∞}^{∞} P(x)dx
$$
```{r}
x <- seq(-5, 5, length = 100)
y <- dnorm(x)
plot(x, y, type = "l")

x <- seq(-5, 1, length = 100)
y <- dnorm(x)
polygon(c(-5, x, 1), c(0, y, 0), col = "gray")
```

```{r}
x <- seq(-5, 5, 0.01)
probDensity <- dnorm(x)
plot(x, probDensity, xlab = "", ylab = "Density", type = "l", main = "PDF of Standard Gaussian")
```

```{r}
cumuDistribution <- pnorm(x)
plot(x, cumuDistribution, xlab = "", ylab = "Cumulative Probability", type = "l", main = "CDF of Standard Gaussian")
```

```{r}
set.seed(42)
hist(rnorm(1000,0,1), main = "Random draws from Std Normal", xlim = c(-4,4))
```

```{r}
y <- 2 * x + rnorm(length(x), 0, 5)
hist(y, prob = TRUE, breaks = 20)
curve(dnorm(x, mean(y), sd(y)), add = TRUE)
```

## Z-score and the empirical rule
The number of standard deviations away from the mean is given by the z-score;
$$
  \frac{x - \mu}{\sigma}.
$$
The empirical rule: $\sigma \in 68.3, 2\sigma \in 95, 3\sigma \in 99.7$.

> Suppose the mean weight of one year olds in kilograms is $\mu = 9.5$, with $\sigma = 1.1$. Without a calculator, estimate the percentage of one year olds that are;
>
i. Less that $8.4$
ii. Between $7.3$ and $11.7$
iii. More than $12.8$

Answers;

i. $P(\mu ± \sigma) = 68$, so $100 - 68 = 32$, the probability of being less that $8.4$ is $32/2 = 16\%$
ii. $P(\mu ± 2\sigma) = 95\%$
iii. $P(\mu ± 3\sigma) = 99.7$, so $100 - 99.7 = 0.3$ and thus percentage that weight more than $12.8$ is $0.3/2 = 0.15\%$

## Confidence Intervals
*KA questions*

> You sample 36 apples from your farms harvest of over 200,000 apples. The mean weight of the sample is 122 grams, with a standard deviation of 40 grams. What is the probability that the mean weight of all 200,000 apples is within 100 and 124 grams?

The question is basically asking this;
$$
\begin{align}
  P(100 ≤ \mu ≤ 124) &= P(\mu \text{ is within 12 of } \bar{x}) \\
    &= P(\bar{x} \text{ is within 12 of } \mu) \\
    &= P(\bar{x} \text{ is within 12 of } \mu_{\bar{x}}) \\
    &=  P(\bar{x} = \mu_{\bar{x}} ± 12). \\
\end{align}
$$
What we have to do is compute how many standard deviations 12 is from $\mu_{\bar{x}}$ and then use the z-table to get the probability;
$$
\begin{align}
  \sigma_{\bar{x}}^{2} &= \frac{\sigma^{2}}{n} \implies \sigma_{\bar{x}} = \frac{\sigma}{\sqrt{n}} ≈ \frac{s}{\sqrt{n}} \\
\text{z-score} &= \frac{x - \mu}{\sigma}.
\end{align}
$$
The z-score in this case is $12 / \sigma_{\bar{x}}$, so the actual question is now, what is?
$$
  P(\bar{x} \text{ is within 1.8 standard deviations of } \mu_{\bar{x}})
$$

```{r}
n <- 36
s <- 40
sd_x <- s / sqrt(n)
z <- 12 / sd_x
pnorm(z)
```
Now to get the range
```{r}
(pnorm(z) - 0.5) * 2
```
There is a 92.81% chance that the population mean $\mu$ is within 12 of the sample mean. Or put another way we can be 92.81% confident that the actual mean is within 100 and 124 grams.

## Bernoulli distribution

The Bernoulli distribution is the simplest case of the binomial distribution.
Consider a poll/survey, were responses are 60% favourable and 40 unfavourable.

Given a population $N$ where the probability of success (1) has probability $p$, that of failure (0) has probability $p - 1$. Then $E[X] = \mu = (1 - p)0 + p = p$, and 
$$
\begin{align}
  \sigma^2 &= (1 - p)(0 - p)^2 + p(1 - p)^2 \\
    &= (1 - p)p^2 + p(1 - 2p + p^2) \\
    &= p^2 - p^3 + p - 2p^2 + p^3 \\
    &= p - p^2 \\
    &= p(1 - p)
\end{align}
$$
In a election with two candidates A and B, the probability of voters going to vote for B is $p$ and for A is $1 - p$. Let 1 be sampling someone going to vote for B and 0 for A.
Suppose we randomly sample 100 voters form a electorate of 20 million; with 57 indicating their going to vote for A and 43 for B.
```{r}
n <- 100
x <- 43 / n
s2 <- (57*(0 - x)^2 + 43*(1 - x)^2) / (n - 1)
s <- sqrt(s2)
s
```
In what interval can we be 95% confident that the true mean $(\mu)$ is within the sample mean $(\bar{x})$ ? From the empirical rule we know that;
$$
  P(\mu = \bar{x} ± 2\sigma_\bar{x}) = 95.4\% = P(p = \bar{x} ± 2(0.05)).
$$
```{r}
sd_x <- s / sqrt(n)
```
Therefore we expect $(0.43 - 0.1) ≤ p ≤ (0.43 + 0.1)$, so 43% will vote for B and 57% for A with a margin of error of 0.1 = 10%.

